### 2014-07-29 Studies/2012_Chicago/Report/Chicago_Fire.tex
Chicago: update report, images, and bib file

\begin{flushright}
\fontsize{20}{24}\selectfont
\bf{NIST Technical Note XXXX}
\end{flushright}

@@

\begingroup
\hypersetup{urlcolor=black}
\href{http://dx.doi.org/10.6028/NIST.TN.XXXX}{http://dx.doi.org/10.6028/NIST.TN.XXXX}
\endgroup
}

@@

\begingroup
\hypersetup{urlcolor=black}
\href{http://dx.doi.org/10.6028/NIST.TN.XXXX}{http://dx.doi.org/10.6028/NIST.TN.XXXX} \\
\endgroup

@@

\large
\flushright{\bf National Institute of Standards and Technology Technical Note XXXX \\
Natl.~Inst.~Stand.~Technol.~Tech.~Note~XX, \pageref{LastPage} pages (August 2014) \\
% http://dx.doi.org/10.6028/NIST.TN.XXXX \\
CODEN: NTNOEF }

@@

The IC made entry to the 1\textsuperscript{st} floor apartment and worked his way to [the rear of the structure]. He opened the [interior door opening into the covered porch on the rear of the house] and noticed heavy fire in the covered porch and rear stairwell area. At 1721 hours, the [E123 captain] and firefighter were on the 2\textsuperscript{nd} floor where they flaked out, charged, and began advancing the hoseline to the rear door of the apartment. The E49 crew had stretched a 2\sfrac{1}{2}-inch hoseline [to the rear of the structure]. The T33 crew set a ground ladder on [the left side of the structure] and TL39 set the aerial to the roof about a third of the way back on [the left side]. A TL39 firefighter went to [the rear of the structure] to check doors. He first went to the basement door which he was unable to force open. Then, he went to the first floor exterior enclosed porch door which was unlocked and he opened it up. He stated that he noticed fire light-up in the stairwell. He kicked in the locked door to the first floor apartment, preceded to walk in, saw no fire then backed out.

@@

\section{Materials}
\label{matl}
While the fires were represented as fuel sources with a constant area (Section~\ref{fire}), from a heat transfer perspective, it is important to define the material properties (density, thermal conductivity, and specific heat) to account for heat transfer and energy storage in the ceiling, walls, and floor. In this study, the material properties of gypsum~\cite{Drysdale:1} were specified on the finished walls and ceilings on the main two floors of the structure, and the material properties of wood~\cite{Incropera:1} were specified on the unfinished portions of the attic and rear porch. The following lines define these materials in the FDS input file:

\begin{lstlisting}
&MATL ID = 'GYPSUM'
    FYI = 'Drysdale' 
    CONDUCTIVITY = 0.17
    SPECIFIC_HEAT = 1.09
    DENSITY = 930. /

&MATL ID = 'WOOD_MAT'
    FYI = 'Incropera and DeWitt, Fundamentals of Heat and Mass Transfer'
    CONDUCTIVITY = 0.12
    SPECIFIC_HEAT = 1.38
    DENSITY = 510. / 
\end{lstlisting}

\section{Ventilation}
\label{Vents}
The simulations in this study account for changes in the ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). Two different ventilation scenarios were considered in this study. The first scenario involves a baseline simulation in which the ventilation times and ventilation areas represent our best understanding of the incident. The second scenario, or alternative simulation, does not include the firefighter opening the rear door to make entry to the structure and the subsequent exterior left wall fire penetration. The times of the ventilation changes are provided in Table~\ref{tab:vents}. Structure leakage has been shown to be important when modeling enclosure fires~\cite{beal2009}. The open front door at the start of the simulation and the vents (opening size and open time) created during the simulation are significantly larger than the total effective area of leakage of a structure of this type. Therefore, any leakage in the structure would have a negligible impact on the fluid mechanics within the structure and is not included in this study. An unknown in Table~\ref{tab:vents} that needed to be estimated was the time when the second floor interior/porch door failed (2~min and 40~s into the simulation). While this time may not be exact, the takeaway from the analysis in the following sections should be impact of the door's failure on the interior conditions of the structure.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=.60\textwidth]{../Figures/smv_exterior_grid}
\caption{Front side of the structure with 10 cm computational mesh.}
\label{fig:geom_grid}
\end{figure}

@@

\bibliography{../../../Bibliography/FDS_refs,../../../Bibliography/FDS_general,chicago}

>>>>

\begin{flushright}
\fontsize{20}{24}\selectfont
\bf{NIST Technical Note 1838}
\end{flushright}

@@

\begingroup
\hypersetup{urlcolor=black}
\href{http://dx.doi.org/10.6028/NIST.TN.1838}{http://dx.doi.org/10.6028/NIST.TN.1838}
\endgroup
}

@@

\begingroup
\hypersetup{urlcolor=black}
\href{http://dx.doi.org/10.6028/NIST.TN.1838}{http://dx.doi.org/10.6028/NIST.TN.1838} \\
\endgroup

@@

\large
\flushright{\bf National Institute of Standards and Technology Technical Note 1838 \\
Natl.~Inst.~Stand.~Technol.~Tech.~Note~1838, \pageref{LastPage} pages (August 2014) \\
% http://dx.doi.org/10.6028/NIST.TN.XXXX \\
CODEN: NTNOEF }

@@

The IC made entry to the 1\textsuperscript{st} floor apartment and worked his way to [the rear of the structure]. He opened the [interior door opening into the rear foyer on the backside of the house] and noticed heavy fire in the covered porch and rear stairwell area. At 1721 hours, the [E123 captain] and firefighter were on the 2\textsuperscript{nd} floor where they flaked out, charged, and began advancing the hoseline to the rear door of the apartment. The E49 crew had stretched a 2\sfrac{1}{2}-inch hoseline [to the rear of the structure]. The T33 crew set a ground ladder on [the left side of the structure] and TL39 set the aerial to the roof about a third of the way back on [the left side]. A TL39 firefighter went to [the rear of the structure] to check doors. He first went to the basement door which he was unable to force open. Then, he went to the first floor exterior enclosed porch door which was unlocked and he opened it up. He stated that he noticed fire light-up in the stairwell. He kicked in the locked door [that connects the 1\textsuperscript{st} floor rear porch to the rear foyer], preceded to walk in, saw no fire then backed out.

@@

\section{Materials}
\label{matl}
While the fires were represented as fuel sources with a constant area (Section~\ref{fire}), from a heat transfer perspective, it is important to define the material properties (density, thermal conductivity, and specific heat) to account for heat transfer and energy storage in the ceiling, walls, and floor. In this study, the material properties of gypsum board~\cite{WAKILI2007} were specified on the finished walls and ceilings on the main two floors of the structure, and the material properties of wood~\cite{Incropera:1} were specified on the unfinished portions of the attic and rear porch. The following lines define these materials in the FDS input file:

\begin{lstlisting}
&MATL ID = 'GYPSUM BOARD'
    FYI = 'Wakili - Journal of Fire Sciences 2007' 
    CONDUCTIVITY = 0.28
    SPECIFIC_HEAT = 1.0
    DENSITY = 810. /

&MATL ID = 'WOOD MAT'
    FYI = 'Incropera and DeWitt, Fundamentals of Heat and Mass Transfer'
    CONDUCTIVITY = 0.12
    SPECIFIC_HEAT = 1.38
    DENSITY = 510. / 
\end{lstlisting}

\section{Ventilation}
\label{Vents}
The simulations in this study account for changes in the ventilation due to a combination of fire department operations (opening doors) and fire acting on the structure (fire breaching walls, breaking windows, etc). Two different ventilation scenarios were considered in this study. The first scenario involves a baseline simulation in which the ventilation times and ventilation areas represent our best understanding of the incident. The second scenario, or alternative simulation, does not include the firefighter opening the rear door to make entry to the structure and the subsequent exterior left wall fire penetration. The times of the ventilation changes are provided in Table~\ref{tab:vents}. Structure leakage has been shown to be important when modeling enclosure fires~\cite{beal2009}. The open front door at the start of the simulation and the vents (opening size and open time) created during the simulation are significantly larger than the total effective area of leakage of a structure of this type. Therefore, any leakage in the structure would have a negligible impact on the fluid mechanics within the structure and is not included in this study. An unknown in Table~\ref{tab:vents} that needed to be estimated was the time when the second floor interior/porch door failed (2~min and 40~s into the simulation). The second estimate was the area of failure. The bottom hinge from the door appeared to be closed throughout the fire due to lack of fire damage and soot on the interior faces the hinge and remaining wood (Fig.~(\ref{fig:door_frame})). It appears that the door collapsed and failed partially open. In the simulations only the upper portion of the door was assumed to fail. While this time and area may not be exact, the takeaway from the analysis in the following sections should be impact of the door's failure on the interior conditions of the structure.

\begin{figure}[!ht]
\centering
\includegraphics[width=.80\textwidth]{../Figures/door_frame_burn}
\caption[Post incident image of rear porch doorframe.]{A post incident image of the second floor porch doorframe that failed taken from inside the structure looking toward the porch. Note that the hinge is still in take and there is under burned, ``clean'' wood at along the bottom portion of the frame.}
\label{fig:door_frame}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}r}
   \scalebox{1.0}{\includegraphics[width=.49\textwidth]{../Figures/smv_exterior_grid}} &
   \scalebox{1.0}{\includegraphics[width=.488\textwidth]{../Figures/smv_exterior_grid_rear}} \\
\end{tabular*}
\caption{Front (left) and rear (right) of the structure with 10 cm computational mesh.}
\label{fig:geom_grid}
\end{figure}

@@

\bibliography{../../../Bibliography/FDS_refs,../../../Bibliography/FDS_general}
